---
title: "ToothGrowth inferential data analysis"
author: "Iain Russell"
date: "04/09/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data and Libraries
 
```{r}
data("ToothGrowth")
library(dplyr)
library(reshape2)
library(ggplot2)
```

## Exploratory data analysis

```{r}
str(ToothGrowth)
summary(ToothGrowth)
head(ToothGrowth)
```

The dataset contains observations of effect of vitamin C on tooth growth in 60 guinea pigs, 30 using orange juice (OJ) delivery method, the other 30 using ascorbic acid (VC) delivery method. 

```{r}
ToothGrowth %>% filter(supp == 'OJ') %>% nrow
ToothGrowth %>% filter(supp == 'VC') %>% nrow

```

Each animal was given a supplement dose of either 0.5,1, or 2 milligrams per day. Therefore each animal in the study is independent.

```{r}
ToothGrowth %>% filter(dose == 0.5) %>% nrow
ToothGrowth %>% filter(dose == 1) %>% nrow
ToothGrowth %>% filter(dose == 2) %>% nrow

```


The boxplot shows that in general the OJ supplement yields greater toothgrowth than the VC supplement, except for the higher dose of 2.

```{r}
ggplot(ToothGrowth, aes(supp, len, fill = supp)) + geom_boxplot() + facet_grid(. ~ dose)

```

## Dose and Supp testing


