---
title: "ToothGrowth inferential data analysis"
author: "Iain Russell"
date: "04/09/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data and Libraries
 
```{r}
data("ToothGrowth")
library(dplyr)
library(reshape2)
library(ggplot2)
```

## Exploratory data analysis

```{r}
str(ToothGrowth)
summary(ToothGrowth)
head(ToothGrowth)
```

The dataset contains observations of effect of vitamin C on tooth growth in 60 guinea pigs, 30 using orange juice (OJ) delivery method, the other 30 using ascorbic acid (VC) delivery method. 

```{r}
ToothGrowth %>% filter(supp == 'OJ') %>% nrow
ToothGrowth %>% filter(supp == 'VC') %>% nrow

```

Each animal was given a supplement dose of either 0.5,1, or 2 milligrams per day. Therefore each animal in the study is independent.

```{r}
ToothGrowth %>% filter(dose == 0.5) %>% nrow
ToothGrowth %>% filter(dose == 1) %>% nrow
ToothGrowth %>% filter(dose == 2) %>% nrow

```


The boxplot shows that in general the OJ supplement yields greater toothgrowth than the VC supplement, except for the higher dose of 2 where they are more similar.

```{r}
g <- ggplot(ToothGrowth, aes(supp, len, fill = supp)) + geom_boxplot() + facet_grid(. ~ dose)
g
```

## Dose and Supp testing

The boxplot indicates differing variances for the OJ and VC supplements, and so t-tests for independent data (not paired) with unequal variances are performed to find out if there are any statistically significant differences between the supplements with each level of Vitamin-C dose.

```{r}
dose05 <- ToothGrowth %>% filter(dose == 0.5)
dose1 <- ToothGrowth %>% filter(dose == 1)
dose2 <- ToothGrowth %>% filter(dose == 2)

summary(dose05); summary(dose1); summary(dose2)
```

```{r}
rbind(
      t.test(len ~ supp, paired = FALSE, var.equal = FALSE, data = dose05)$conf,
      t.test(len ~ supp, paired = FALSE, var.equal = FALSE, data = dose1)$conf,
      t.test(len ~ supp, paired = FALSE, var.equal = FALSE, data = dose2)$conf
)
```

The 95% confidence interval for dose levels 0.5 and 1 (mg/day) are entirely above zero, indicating that for these dose levels the OJ supplement produces greater tooth growth than the VC supplement.  

At the higher dose level 2 (mg/day) the interval is either side of zero indicating that we cannot rule out (with 95% confidence), that there is no effect on tooth growth between OJ and VC supplement/delivery methods.

## Conclusion

Under the assumptions that the observations were made on 60 different/independent animals, and that the sample size of 20 is sufficient to assume an underlying normal distribution, the statistical evidence suggests that the OJ delivery method is more effective at producing greater tooth growth at low dose levels (0.5, 1). However there is no conclusive evidence for a difference between delivery methods at higher dose levels of vitamin-C.

